### 2015-06-23 Tex_AXYB/paper.tex
Revised First Two Sections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{cite}      % Written by Donald Arseneau
                        % V1.6 and later of IEEEtran pre-defines the format
                        % of the cite.sty package \cite{} output to follow
                        % that of IEEE. Loading the cite package will
                        % result in citation numbers being automatically
                        % sorted and properly "ranged". i.e.,
                        % [1], [9], [2], [7], [5], [6]
                        % (without using cite.sty)
                        % will become:
                        % [1], [2], [5]--[7], [9] (using cite.sty)
                        % cite.sty's \cite will automatically add leading
                        % space, if needed. Use cite.sty's noadjust option
                        % (cite.sty V3.8 and later) if you want to turn this
                        % off. cite.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/

\usepackage[dvips]{graphicx}  % Written by David Carlisle and Sebastian Rahtz
                        % Required if you want graphics, photos, etc.
                        % graphicx.sty is already installed on most LaTeX
                        % systems. The latest version and documentation can
                        % be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
                        % Another good source of documentation is "Using
                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
                        % which can be found as esplatex.ps and epslatex.pdf
                        % at: http://www.ctan.org/tex-archive/info/

%\usepackage{amsmath}   % From the American Mathematical Society
                        % A popular package that provides many helpful commands
                        % for dealing with mathematics. Note that the AMSmath
                        % package sets \interdisplaylinepenalty to 10000 thus
                        % preventing page breaks from occurring within multiline
                        % equations. Use:
\usepackage{multirow}
%\usepackage[left=0.71in,top=0.94in,right=0.71in,bottom=1.18in]{geometry}
%\setlength{\columnsep}{0.24in}
% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
\usepackage{mathtools}   % loads »amsmath«
\usepackage{amssymb,amsmath}
\usepackage{amsfonts}
\usepackage{color}

\title{\LARGE \bf
Simultaneous Coordinate Calibrations by Solving the AX=YB Problem without Correspondence*
}

\author{Haiyuan Li$^{1}$, Qianli Ma$^{2}$, Tianmiao Wang$^{1}$ and Gregory Chirikjian$^{2}*$% <-this % stops a space
\thanks{*This work is partially
supported by NSF Grant RI-Medium: \#IIS-1162095}% <-this % stops a space
\thanks{$^{1}$Haiyuan Li and Tianmiao Wang is with the School of Mechanical Engineering and Automation, Beihang University,
        ,Beijing 100191, China
        {\tt\small lihaiyuan@buaa.edu.cn}}%
\thanks{$^{2}$Qianli Ma and Gregory Chirikjian with the Department of Mechanical Engineering, Johns Hopkins University,
        Baltimore, MD 21218, USA
        {\tt\small gregc@jhu.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

In image-guided system, relationships of hand-eye ($X$) and robot-world ($Y$) coordinates have to be calculated and simultaneous solution is useful in sensor calibration problem. Due to asynchrony of sensors timing, the correspondence between $A$ and $B$ is unknown. A probabilistic method is presented to solve the
homogeneous matrix equations without a priori knowledge of the correspondence. Using Euclidean-Group invariants, an exact solution can be found. We illustrate the calculation in numeric simulation including various numbers of
robot movements. The results show the efficiency and robustness of the proposed simultaneous
method.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
% no \PARstart

Image-guided system has been widely used in robotics such as robot assisted surgery, (more examples). Sensors such as a camera, a laser scanner or an ultrasound probe are usually mounted on the distal end of a robotic manipulator. For a typical ``hand-eye'' system as described above, the relative transformation of the sensor with respect to the end-effector should be accurately calibrated, and it is often characterized as the well known $\textbf{AX=XB}$ problem. A variation of this problem is characterized as the $\textbf{AX=YB}$ problem, where both the hand-eye transformation and the pose of the robot base with respect to the world frame need to be calibrated. In a typical environment setup, the relationship among the sensor frame, robot frame and world frame is variant and the uncertainties exist. Therefore, simultaneous coordinate calibrations have to be determined frequently in order to enable the robots to respond to dynamic environments.

In the $\textbf{AX=YB}$ problem, $As$ and $Bs$ can be respectively obtained via different sensors. The data streams can be in an asynchronous fashion due to the different working frequencies of the sensors. The asynchrony causes a shift between the two streams of data which damages the correspondence between $As$ and $Bs$. In this paper, a novel method is presented to solve for an $X$ and $Y$ without the need to know a priori knowledge of the correspondence between $As$ and $Bs$.

\begin{center}
\begin{figure}[htbp]
\centering
\includegraphics[width=3.2in]{fig1.eps}
\caption{
(1) The hand-eye and robot-world calibration problem which is formulated as AX=YB (The universal robot as shown in the picture is owned by professor Emad Boctor in). (2) The hand-eye calibration problem which is formulated as AX=XB.
}
\label{fig1}
\end{figure}
\end{center}
The hand-eye calibration problem can be modeled as the $\textbf{AX = XB}$, where $A$ and $B$ are the homogeneous transformation matrices describing  the relative motions of end-effector and the  sensor respectively. As shown in Fig.~\ref{fig1}, $A = A^{i}(A^{i+1})^{-1}$ and  $B = B^{i}(B^{i+1})^{-1}$. Given multiple pairs of $\{A,B \}$ with correspondence, many methods have been proposed to solve for $X$. To the best of the authors' knowledge, Shiu \cite{Shiu1989} and Tsai \cite{Tsai1989} are the first to solve the $AX = XB$ sensor calibration problem. The other methods include but are not limited to the quaternion, dual quaternion, screw theory, Lie group theory convex optimization and gradient descent methods \cite{Wang1992,Park1994,Horaud1995,Daniilidis1999,Fassi2005,Zhao2011,Ackerman2014a}. All of the methods above assume a prior knowledge of exact correspondence between $A_i$ and $B_i$. For data streams $\{A_i\}$ and $\{B_i\}$ that are asynchronous, several methods have been proposed in the literature to solve for $X$ using data without correspondence.  Theses methods assume that there is exact knowledge of the $As$ and $Bs$ correspondence \cite{Ackerman2013a,Ackerman2013,Ackerman2014}.

Simultaneous estimation of the hand-eye and robot-world transformations has been viewed as the $\textbf{AX=YB}$ problem. As shown in Fig.~\ref{fig1}, $Y$ is the transformation from the robot base to the world frame, $A$ denotes the pose of the sensor in the world frame and $B$ is the transformation from the the end-effector to its fixed base. The $A$ and $B$ in $\textbf{AX=YB}$ are different from those in $\textbf{AX=XB}$ where the former uses absolute transformations and the latter uses relative transformations. This problem has been solved by many different methods such as kronecker product, quaternion, dual quaternion, and nonlinear optimization methods \cite{Zhuang1994,dornaika1998simultaneous,Hirsh2001, ernst2012non,strobl2006optimal,Li2010,Shah2013,Heller2014}. Simulatneous calibration of $X$ and $Y$ can be problematic in that all the methods above assume exact correspondence between $\{A_i\}$ and $\{B_j\}$, which is not the case in the real world. Another similar problem involves the calibration of multiple robots in terms of hand-eye, tool-flange and robot-robot system, and it is formulated as the $\textbf{AXB=YCZ}$ problem \cite{Wang2014}. Simultaneous solution for $X$ and $Y$ in $\textbf{AX=YB}$ problem is an challenging issue. In the above methods, the correspondence between $A$ and $B$ is known a prior. In this paper, we focus on the $\textbf{AX=YB}$ problem which does not require a priori knowledge of the correspondence of the data streams.

The rest of the paper is organized as follows. In Section
\ref{sect2}, a novel probabilistic method is presented to solve for $X$ and $Y$. In Section \ref{sect3}, an algorithm involving both correlation theorem and Euclidean group invariants is proposed to recover the correspondence between $\{A_i\}$ and $\{B_j\}$. The simulation results which deal with noisy data without correspondence are illustrated in Section \ref{sect4}. Finally, conclusions are drawn based on the numerical results and possible future works are pointed out.

\section{Solving AX=YB using a probabilistic method on motion groups}
\label{sect2}
In this section, a brief introduction to the concepts of probability density function on the special Eucledian group $SE(3)$ is presented and the probabilistic representation of $AX=YB$ are derived.

Any rigid transformation matrix can be viewed as a group element of $SE(3)$ :
\begin{equation}\label{equ0}
    H(R,t)=\left(
             \begin{array}{cc}
               R & t \\
               0^{T} & 1 \\
             \end{array}
           \right) \in SE(3)
\end{equation}
where $SO(3)$ denotes the special orthogonal group, $t \in \mathbb{R}^3 $ is translational vector and $H$ is symbol for group element.

Given a large set of pairs $(A_{i},B_{i})\in SE(3)\times SE(3)$ where $i=1,\cdots,n$, the following equation is true if the correspondence is known as a priori:

\begin{equation}\label{equ1}
A_{i}X=YB_{i}.
\end{equation}
For a matrix $H \in SE(3)$, a Dirac delta function $\delta(H)$ is defined to be finite only at the identity and zero elsewhere:

\begin{equation}\label{equ2}
\delta{(H)}=
\left\{
\begin{array}{ll}
+\infty, & H=I \\
0, & H \neq I
\end{array}
\right.
\end{equation}
It also satisfies the identity constraint as:

\begin{equation}\label{equ3}
\int_{SE(3)}\delta{(H)}dH=1.
\end{equation}

A shifted Dirac delta function can be defined as $\delta_{A}(H)=\delta{(A^{-1}H)}$. Given $K,H \in SE(3)$ and two well-defined functions $f_{1}, f_{2} \in \left(L^1 \cap L^2 \right)(SE(3))$, their convolution on $SE(3)$ is defined as:

\begin{equation}\label{equ4}
(f_{1}\ast f_{2})(H)=\int_{SE(3)}f_{1}(K)f_{2}(K^{-1}\circ H)dK.
\end{equation}
Employing the properties of $\delta$ function, it is straightforward to see that:

\begin{equation}\label{equ5}
(f\ast \delta)(H)=\int_{SE(3)}f(K)\delta(K^{-1}\circ H)dK=f(H).
\end{equation}
Therefore, for each $A_{i}$ and $B_{i}$, the following equations can be obtained:

\begin{IEEEeqnarray}{l}\label{equ6}
(\delta_{A_{i}}\ast \delta_{X})(H)=\delta(A_{i}^{-1} H X^{-1}) \IEEEyessubnumber
\\*
(\delta_{Y}\ast \delta_{B_{i}})(H)=\delta(Y^{-1} H B_{i}^{-1}). \IEEEyessubnumber
\end{IEEEeqnarray}
Using Eq.($\ref{equ1}$), the above two equations can be combined into a single equation as:

\begin{equation}\label{equ7}
(\delta_{A_{i}}\ast \delta_{X})(H)=(\delta_{Y}\ast \delta_{B_{i}})(H)
\end{equation}

Define the probability density function of $\{A_i\}$ and $\{B_i\}$ as:
\begin{IEEEeqnarray}{l}\label{equ8}
f_{A}(H)=\frac{1}{n}\sum_{i=1}^{n} \delta(A_{i}^{-1}H) \IEEEyessubnumber
\\*
f_{B}(H)=\frac{1}{n}\sum_{i=1}^{n} \delta(B_{i}^{-1}H) \IEEEyessubnumber
\end{IEEEeqnarray}
Using the distributivity of convolution, Eq.(\ref{equ8}) can be substituted into Eq.(\ref{equ7}) to get:
\begin{equation}\label{equ9}
(f_{A_{i}}\ast \delta_{X})(g)=(\delta_{Y}\ast f_{B_{i}})(g)
\end{equation}

For each of the data stream $\{A_i\}$ and $\{B_i\}$, small relative motions are calculated between consecutive reference frames. Take $\{A_i\}$ for example, one metric of the distance between $A_i$ and $A_{i+1}$ can be defined as:

\begin{eqnarray}\label{equ10}
d^{2}(A_{i},A_{i+1})=\parallel \Delta A \parallel_{W}^{2} = trace[(\Delta A)W(\Delta A)^{T}] = \epsilon,\nonumber \\
\end{eqnarray}
where $\Delta A = A_{i}-A_{i+1}$ and $0 < \epsilon \ll 1 $.

The convolution of two ``highly focused" probability density functions (PDF) have some interesting properties that can be used to solve for $X$. In particular, define the mean $M$ and covariance $\Sigma$ of a probability density function on $SE(3)$ as:

\begin{IEEEeqnarray}{l}\label{equ11}
\int_{SE(3)}log(M^{-1}H))f(H)dH=0 \IEEEyessubnumber
\\*
\Sigma = \int_{SE(3)}log^{\vee}(M^{-1}H)[log^{\vee}(M^{-1)}H)]^{T}f(H)dH. \IEEEyessubnumber
\end{IEEEeqnarray}
Then the corresponding discrete version is:

\begin{IEEEeqnarray}{l}\label{equ12}
\sum_{i=1}^{n}log(M^{-1}H))=0 \IEEEyessubnumber
\\*
\Sigma = \sum_{i=1}^{n}log^{\vee}(M^{-1}H)[log^{\vee}(M^{-1)}H)]^{T}. \IEEEyessubnumber
\end{IEEEeqnarray}

Given $\{A_i\}$ where the cloud of frames ${A_{i}}$ clustering around $M_{A}$, an iterative formula can be used for computing $M_{A}$ \cite{Wang2008} as:

\begin{equation}\label{equ13}
    ^{k+1}M_{A} = ^{k}M_{A} \circ exp[\frac{1}{n}\sum_{i=1}^{n}log(^{k}M_{A}^{-1}\circ A_{i})]
\end{equation}

An initial estimate of the iterative procedure can be chosen as $^{0}M_{A}=\frac{1}{n}\sum_{i=1}^{n}log(A_{i})$, then a local minimum of $M_A$ is obtained by solving a nonlinear optimization problem with the cost function being $|| \sum_{i=1}^{n}log(M_{A}^{-1}A_{i}) ||^{2}$. A similar procedure can be used to compute $M_B$. $\Sigma_A$ and $\Sigma_B$ are then straight forward to compute given known $M_A$ and $M_B$.

The mean and covariance for the convolution $(f_{1} \ast f_{2})(g)$ of two highly focused functions $f_{1}$ and $f_{2}$ are calculated as in \cite{Wang2008}:

\begin{IEEEeqnarray}{l}\label{equ14}
M_{1 \ast 2} = M_{1}M_{2} \IEEEyessubnumber
\\*
\Sigma_{1 \ast 2} = Ad(M_{2}^{-1})\Sigma_{1}Ad^{T}(M_{2}^{-1}) + \Sigma_{2}. \IEEEyessubnumber
\end{IEEEeqnarray}
where

$$Ad(H)=\left(
               \begin{array}{cc}
                 R & O \\
                 \hat{t}R & R \\
               \end{array}
             \right).$$

Because $X$ and $Y$ are constant, their corresponding PDF will be $\delta_{X}(g)$ and $\delta_{Y}(g)$, of which the mean and covariance are $M_{X} = X$, $\Sigma_{X} = \mathbb{O}_{6 \times 6}$ and $M_{Y} = Y$, $\Sigma_{Y} = \mathbb{O}_{6 \times 6}$, respectively. Therefore, the following equations can be obtained using Eq.(\ref{equ14}):

\begin{IEEEeqnarray}{l}
M_{A} X = Y M_{B} \IEEEyessubnumber\label{equ15a}
\\*
Ad(X^{-1})\Sigma_{A}Ad^{T}(X^{-1}) = \Sigma_{B}. \IEEEyessubnumber\label{equ15b}
\end{IEEEeqnarray}

To solve the above equations, Eq.(\ref{equ15a}) is decomposed into a rotational equation and a translational equation as below:

\begin{IEEEeqnarray}{l}
R_{M_{A}} R_{X} = R_{Y} R_{M_{B}} \IEEEyessubnumber\label{equ16a}
\\*
R_{M_{A}} t_{X} + t_{M_{A}}= R_{Y} t_{M_{B}} + t_{Y}. \IEEEyessubnumber\label{equ16b}
\end{IEEEeqnarray}
$\Sigma_{A}$ and $\Sigma_{B}$ can be decomposed into blocks as
$\left(\begin{array}{cc}
       \Sigma_{A}^{1} & \Sigma_{A}^{2} \\
       \Sigma_{A}^{3} & \Sigma_{A}^{4} \\
       \end{array}
       \right)$
and
$\left(\begin{array}{cc}
       \Sigma_{B}^{1} & \Sigma_{B}^{2} \\
       \Sigma_{B}^{3} & \Sigma_{B}^{4} \\
       \end{array}
       \right)$, respectively. Knowing that
$ X^{-1} = \left(\begin{array}{cc}
       R_{X}^{T} & -R_{X}^{T}t_{X}  \\
       0 & 1 \\
       \end{array}
       \right)$,
then the first two blocks of Eq.(\ref{equ15b}) can be written as follows:

\begin{IEEEeqnarray}{l}
\Sigma_{M_{B}}^{1} = R_{X}^{T}\Sigma_{M_{A}}^{1} R_{X} \IEEEyessubnumber\label{equ17a}
\\*
\Sigma_{M_{B}}^{2} = R_{X}^{T}\Sigma_{M_{A}}^{1} R_{X} (\widehat{R_{X}^{T}t_{X}}) + R_{X}^{T}\Sigma_{M_{A}}^{2}R_{X}.
 \IEEEyessubnumber\label{equ17b}
\end{IEEEeqnarray}
Because Eq.(\ref{equ17a}) is a similarity transformation between $\Sigma_{M_{B}}^{1}$ and $\Sigma_{M_{A}}^{1}$, they share the same eigenvalues and can be eigendecomposed into   $\Sigma_{M_{A}}^{1}=Q_{M_{A}}\Lambda Q_{M_{A}}^{T}$ and $\Sigma_{M_{B}}^{1}=Q_{M_{B}}\Lambda Q_{M_{B}}^{T}$ where $\Lambda$ is a diagonal matrix whose diagonal elements are the eigenvalues of $\Sigma_{M_{A}}^{1}$ ($\Sigma_{M_{B}}^{1}$), and $Q_{M_{A}}$ ($Q_{M_{B}}$) is a square matrix whose columns are the corresponding eigenvectors. Following equation is obtained after substitution into Eq.(\ref{equ17a}):

\begin{equation}\label{equ18}
    \Lambda = (Q_{M_{A}}^{T}R_{X}^{T}Q_{M_{B}}) \Lambda (Q_{M_{B}}^{T}R_{X}Q_{M_{A}})= P \Lambda P^{T}
\end{equation}
where $P=Q_{M_{A}}^{T}R_{X}Q_{M_{B}}$. If $Q_{M_{A}}$ and $Q_{M_{B}}$ are further constrained to be rotation matrices, then a rotation matrix $P$ that satisfies Eq.(\ref{equ18}) can be:

\begin{equation}\label{equ19}
\begin{split}
{\cal P} = \left\{ \left(\begin{array}{ccc}
1\,\, & 0 \,\, & 0 \,\, \\
0 & 1 & 0\\
0 & 0 & 1 \end{array}\right) , \left(\begin{array}{ccc}
-1 & 0 & 0 \\
0 & -1 & 0\\
0 & 0 & 1 \end{array}\right), \right.\\
\left. \left(\begin{array}{ccc}
-1 & 0 & 0 \\
0 & 1 & 0\\
0 & 0 & -1 \end{array}\right), \left(\begin{array}{ccc}
1 & 0 & 0 \\
0 & -1 & 0\\
0 & 0 & -1 \end{array}\right)\right\}.
\end{split} 
\end{equation}

Therefore, there are eight candidates of $R_{X}$ which can be calculated via $R_{X}=Q_{M_{A}}PQ_{M_{B}}^T$. Then the corresponding $t_{X}$ can be obtained  from Eq.(\ref{equ17b}). Given known $X$, $Y$ can be solved for by $Y = M_A^{-1}XM_B^{-1}$. At last, eight candidate pairs of  $\{X_{k},Y_{k}\}$ can be obtained as:

\begin{equation}\label{equ20}
\begin{array}{cc}
X_{k}= \left( \begin{array}{cc}
       R_{X_k} & t_{X_k} \\
       \mathbf{0}^{T} & \mathbf{1}\\
\end{array} \right),&
Y_{k}= \left( \begin{array}{cc}
       R_{Y_k} & t_{Y_k} \\
       \mathbf{0}^{T} & \mathbf{1}\\
\end{array} \right)
\end{array}
\end{equation}
where $k = 1,2,...,8$.

The problem then becomes selecting the best pair of $\{X_k, Y_k\}$ from the eight candidates. 
Based on the screw theory (references?), it is known that a homogeneous transformation $H$ can be expressed by the four screw parameters $(\theta,d,\mathbf{n},\mathbf{p})$ as:

\begin{equation}\label{equ21}
H = \left( \begin{array}{cc}
       e^{\theta \hat{\mathbf{n}}} & (\mathbf{I}_{3} - e^{\theta \hat{\mathbf{n}}})\mathbf{p} + d\mathbf{n} \\
       \mathbf{0}^{T} & \mathbf{1}
\end{array} \right)
\end{equation}
where $\theta$, $d$, $\mathbf{n}$ and $\mathbf{p}$ are ?.

Moreover, $AX_k = Y_k B$ can be written as $AX_k=X_k(X_k^{-1}Y_kB)$. If define $B^k = X_k^{-1}Y_kB$, then we will have $AX_k=X_kB^k$. There exist two Euclidean-Group invariant relationships for each pair of $(A_{i},B_{i}^k)( i = 1,\cdots,n; k=1,\dots,8)$ as follows:

\begin{equation}\label{equ22}
    \theta_{A_{i}}=\theta_{B_{i}^{k}}, d_{A_{i}}=d_{B_{i}^{k}}
\end{equation}

Among the four pairs $(X_{k},Y_{k})$, one can find an optimal solution which  minimizes the cost function defined as:

\begin{equation}\label{equ23}
    (X,Y) = \mathop{\mathbf{arg}min}_{(X_{k},Y_{k})}\frac{1}{n} \sum_{i=1}^{n} (\parallel \theta_{A_{i}}-\theta_{B_{i}^{k}} \parallel + \parallel d_{A_{i}}-d_{B_{i}^{k}} \parallel)
\end{equation}

\section{Solution with unknown correspondence between $A_{i}$ and $B_{i}^{k}$}
\label{sect3}

In most cases, the homogeneous transformations with $As$ and $Bs$ are given based on the data from different sensors. Due to asynchronous timing of the measurement transmissions, the correspondences between $A_{i}$ and $B_{i}^{k}$ is unknown. The advantage of the above probabilistic solution lies on that $X$ and $Y$ can be calculated even if without any a priori knowledge of the correspondence. However, there are still eight possible candidate results $(X_{i},Y_{i})$. Using Euclidean-Group invariants, it is straightforward to determine which pair is the correct one if the correspondence between $A_{i}$ and $B_{i}^{k}$ can be known.

The Discrete Fourier Transform (DFT) decomposes a time-domain signal into its constituent frequencies. The input is a finite list of equally spaced samples of a function. Given a discrete signal consisting of a sequence of $N$ complex numbers $x_{0},x_{1},\cdots,x_{N-1}$, the DFT is denoted by $X_{\kappa} = \mathcal{F}{x_{n}}$

\begin{equation}\label{equ21}
    X_{\kappa} = \sum_{n=0}^{N-1}x_{n}\cdot exp(-i\frac{2\pi}{N}n\kappa)
\end{equation}
The Inverse Discrete Fourier transform (IDFT) denoted by

\begin{equation}\label{equ22}
    x_{n} = \frac{1}{N}\sum_{n=0}^{N-1}X_{\kappa}\cdot exp(i\frac{2\pi}{N}n\kappa)
\end{equation}

The discrete convolution of two sequences $f_{n}$ and $g_{n}$  are defined

\begin{equation}\label{equ23}
    (f \ast g)(\tau)=\sum_{i=0}^{N}f(t_{i})g(t_{i}-\tau)
\end{equation}

In convolution theorem, the Fourier transform of a convolution is the product of the Fourier transforms, namely,

\begin{equation}\label{equ24}
    f \ast g = \mathcal{F}^{-1} [\mathcal{F}(f) \cdot \mathcal{F}(g)]
\end{equation}

The correlation theorem indicates that the correlation function, $Corr(f,g)$, will have a large value at a shift vector if the two sequences $f$ and $g$ contain similar features. The correlation can be obtained based on the convolution theorem. The DFT of the correlation $Corr(f,g)$ is equal to the product of the DFT of a sequence $f_{n}$ and the complex conjugate $\mathcal{F}^{*}$ of the DFT of the other sequence $g_n$.

\begin{equation}\label{equ25}
    Corr(f,g)=f \star g = \mathcal{F}^{-1}[\mathcal{F}(f) \cdot (\mathcal{F}(g))^{*}]
\end{equation}

Compared with the standard time-domain convolution algorithm, the complexity of the convolution by multiplication in the frequency domain is significantly reduced with the help of the convolution theorem and the fast Fourier transform (FFT).

There are two sequences $\theta_{A_{i}}$ and $\theta_{B_{i}^{k}}$ from each pair $(A_{i},B_{i}^{k})$. For homogeneous transformations from which the range of $\theta$ can vary, two sequences $\theta_{Ai}$ and $\theta_{B_{i}^{k}}$ can be first normalized.

\begin{equation}\label{equ26}
    \theta_{1}=\frac{(\theta_{A_{i}}-\mu_{A_{i}})}{\sigma_{A_{i}}}, \theta_{2}=\frac{(\theta_{B_{i}^{k}}-\mu_{B_{i}^{k}})}{\sigma_{B_{i}^{k}}}
\end{equation}

where $\mu_{A_{i}}(\mu_{B_{i}^{k}})$ is the average of $\theta_{A_{i}}(\theta_{B_{i}^{k}})$ and $\sigma_{A_{i}}(\sigma_{B_{i}^{k}})$ is the standard deviation.

Here, the correlation function $Corr(\theta_{1},\theta_{2})$  is the function of the time sequence index $(n)$ which describes the probability that these two sequences are separated by this particular unit. The location of the function maximum indicates the amount of shift, $\tau_{shift}$, between the two sequence $\theta_{A_{i}}$ and $\theta_{B_{i}^{k}}$.

\begin{equation}\label{equ27}
    \tau_{shift} = \mathop{\mathbf{arg}max}_{index}(Corr(\theta_{1},\theta_{2}))
\end{equation}

Therefore, the correspondence between the two sequences can be found. The data of $\theta_{A_{i}}$ or $d_{A_{i}}$ are shifted by $-\tau_{shift}$ to obtain a sequence of new pairs $(\theta_{A_{i}}(i+\tau_{shift}),\theta_{B_{i}^{k}})$ and $(d_{A_{i}}(i+\tau_{shift}),d_{B_{i}^{k}})$, $max(i,i+\tau_{shift})\leq i \leq min(i,i+\tau_{shift})$. The data stream can be shifted to reach correspondence once the shift is found and the correct solution can also be found by minimizing the absolute deviations based on Euclidean-Group invariants relations using the method in Section \ref{sect2}.

\section{Simulation Studies}
\label{sect4}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=2.6in]{fig3.eps}
\caption{
$Bs$ are randomly generated using  $B_i=B_{init}exp(\widehat{\mathbf{x}})$,$\mathbf{x}=(x_1,\dots,x_6)^T $ \ $x_i \sim \mathcal{N}(0,0.1)$.
}
\label{fig2}
\end{figure}
\end{center}

\begin{figure}
\centering
\includegraphics[width=2.6in,bb=84 202 501 581]{fig2.eps}
\caption{
$As$ are calculated using $AX=YB$. $X$ and $Y$ are assumed.
}
\label{fig3}
\end{figure}

In the numerical experiments in this section, the rotational and translational error for $X$ and $Y$ are measured as  $Error(R_X) = \parallel log^{\vee} (R_{X_{Solved}}^{T}R_{X_{true}})\parallel$, $Error(t_X) = \parallel (t_{X_{Solved}}-t_{X_{true}})\parallel$, $Error(R_Y) = \parallel log^{\vee} (R_{Y_{Solved}}^{T}R_{Y_{true}})\parallel$ and $Error(t_Y) = \parallel (t_{Y_{Solved}}-t_{Y_{true}})\parallel$ respectively.

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig4.eps}
\caption{
The cross correlation of data streams of $(A_{i},B_{i}^{k})$ respectively.
}
\label{fig4}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig5.eps}
\caption{
Calculated rotational and translational deviation of $X$ and $Y$ solved using the data in Fig.~\ref{fig2} and Fig.~\ref{fig3}.
}
\label{fig5}
\end{figure}
\end{center}

$B_{i}$ are generated randomly closely around $B_{init}$ using $B_i = B_{init} exp(\mathbf{\widehat{x}})$ and $i$ pose measurements were employed for generating $i$ $A_{i}$ by $A_{i} = YB_{i}X^{-1}$ as shown in Fig.~\ref{fig2} in which a example shows $A$ and $B$ distribution. As a result by applying the above probabilistic method, 8 sequences $(\theta_{A_{i}},\theta_{B_{i}^{k}})$ and $(d_{A_{i}},d_{B_{i}^{k}})$ $ (i=5,\cdots, 100, k=1,\dots,8)$ can be obtained respectively.

If the data streams of $As$ were shifted by $m$ units compared to the data stream $Bs$. The maximum of cross correlation can be used to find the corresponding shift, which is $-m$ shown in Fig.~\ref{fig4}, representing the data stream of $B_{i}^{k}$ has been shifted by -m units with respect to ${A_{i}}$. Therefore, we shift the data stream inversely to recover the correspondence for finding a correct solution satisfying Euclidean-Group invariants.

Using the minimum sum of $||\theta_{A_{i}} - \theta_{B_{i}^{k}}||$ and $||d_{A_{i}} - d_{B_{i}^{k}}||$, we can find the ${B_{i}^{k}}(k=3)$ corresponding to the least sum of errors and then, only a $(X_k,Y_k)$ is the desired solution. In Fig.~\ref{fig5}, as the number of $(A,B)$ pairs increase, the errors of translation and rotations are reduced but when the number comes to a certain value, the errors cannot be reduced furthermore.

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig6.eps}
\caption{
$Error(R_X)$ distribution as the $As$ and $Bs$ spread ($\sigma(x_i) = 0.1, 0.2, 0.4, 0.6, 0.8, 1, 2$ as shown in Fig.~\ref{fig2})
}
\label{fig6}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig7.eps}
\caption{
$Error(R_Y)$ distribution as the $As$ and $Bs$ spread ($\sigma(x_i) = 0.1, 0.2, 0.4, 0.6, 0.8, 1, 2$ as shown in Fig.~\ref{fig2})
}
\label{fig7}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig8.eps}
\caption{
$Error(t_X)$ distribution as the $As$ and $Bs$ spread ($\sigma(x_i) = 0.1, 0.2, 0.4, 0.6, 0.8, 1, 2$ as shown in Fig.~\ref{fig2})
}
\label{fig8}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=3.2in]{fig9.eps}
\caption{
$Error(t_Y)$ distribution as the $As$ and $Bs$ spread ($\sigma(x_i) = 0.1, 0.2, 0.4, 0.6, 0.8, 1, 2$ as shown in Fig.~\ref{fig2})
}
\label{fig9}
\end{figure}
\end{center}

In the generation of $B_i = B_{init} exp(\mathbf{\widehat{x}})$, each element $x_j$ of $\mathbf{x} = (x_1,x_2,x_3,x_4,x_5,x_6)^T $ is Gaussian with $N \sim (\mu,\sigma)$. Small disturbances are exerted to the $B_i$ to make the noisy $B_i^{noise} = B_i exp(\mathbf{\widehat{x}}_{noise})$, where each of Lie Algebra element of $\mathbf{x}$ is Gaussian distribution $N \sim (\mu_{noise},\sigma_{noise})$. In Fig.~\ref{fig8},~\ref{fig9},~\ref{fig10} and~\ref{fig11}, $\sigma_{noise}$ is 0.005. As $\sigma$ varies from $0.1$ to $2$, the errors of $R_X$, $R_Y$, $t_X$, and $t_Y$ are reduced as shown in the box-and-whisker plot. There are several outliers not included between the whiskers. The median data can be used as the final solved $X$ and $Y$.

\begin{center}
\begin{figure}
\centering
\includegraphics[width=2.6in]{fig10.eps}
\caption{
The distribution of solved $X$.
}
\label{fig10}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\centering
\includegraphics[width=2.6in]{fig11.eps}
\caption{
The distribution of solved $Y$.
}
\label{fig11}
\end{figure}
\end{center}

\section{Conclusions}
\label{sect5}

In this paper, we developed a probabilistic approach to simultaneously obtain  $X$ and $Y$ in $AX=YB$ sensor calibration problem. Without a prior knowledge of the correspondence between $A$ and $B$, in the algorithm the probability theory in Lie group is used to constrain the solution of $X$ and $Y$ to eight candidates. As for the shifted data stream of $A$ and $B$, using the correlation theorem with Euclidean group invariants,the correspondence is recovered to determine the correct solution from eight candidates. In numeric simulation, the method perform well with different data samples.
%\section{Conclusion}
% conference papers do not normally have an appendix

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

\section*{ACKNOWLEDGMENT}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,bibIEEEfull}
%\begin{thebibliography}{1}
%
%
%\end{thebibliography}

\end{document}